---
title: "Interactive Graphs With Plotly And Ggiraph Packages"
date: 2017-01-26
---

Imagine that you have a plot like this and want to make it interactive (e.g. show exact values on hover or print the model name).

```{r}
suppressPackageStartupMessages(library(ggplot2))

p <- ggplot(mpg, aes(displ, hwy, color=class)) + 
  geom_point() +
  labs(title="Smaller car = better MPG",
       subtitle="Fuel economy data from 1999 and 2008 for 38 popular models of car",
       caption = "adapted from http://fueleconomy.gov",
       x="engine displacement (l)",
       y="highway (miles per gallon)")

print(p)
```

## Plotly

Getting `plotly` of this graph is just one-liner. From now all ggplots can be easily made interactive.

```{r}
suppressPackageStartupMessages(library(plotly))

ggplotly(p)
```

And if you need some other information to be printed (like `model`), add it to aesthetics:

```{r}
p2 <- ggplot(mpg, aes(x=displ, y=hwy, color=class, model=model)) + 
  geom_point() +
  labs(title="Smaller car = better MPG",
       subtitle="Fuel economy data from 1999 and 2008 for 38 popular models of car",
       caption = "adapted from http://fueleconomy.gov",
       x="engine displacement (l)",
       y="highway (miles per gallon)")

ggplotly(p2)
```

If you want to show only `model` label:

```{r}
p2 <- ggplot(mpg, aes(x=displ, y=hwy, color=class, text=model)) + 
  geom_point() +
  labs(title="Smaller car = better MPG",
       subtitle="Fuel economy data from 1999 and 2008 for 38 popular models of car",
       caption = "adapted from http://fueleconomy.gov",
       x="engine displacement (l)",
       y="highway (miles per gallon)")

# the only tooltip is now `text`
ggplotly(p2, tooltip = "text")
```


Plotly is able to create graphs with tens of thousands data points.

## Ggiraph

Another way to create analogical plots is `ggiraph` package.

```{r}
suppressPackageStartupMessages(library(ggiraph))
p3 <- ggplot(mpg, aes(x=displ, y=hwy, color=class, tooltip=model)) + 
  geom_point_interactive(size=I(2)) + 
  labs(title="Smaller car = better MPG",
       subtitle="Fuel economy data from 1999 and 2008 for 38 popular models of car",
       caption = "adapted from http://fueleconomy.gov",
       x="engine displacement (l)",
       y="highway (miles per gallon)")

ggiraph(code = print(p3), selection_type = "single")
```

It not only looks great but it also works with R/Shiny (e.g. clicking on a dot can be used as input), see [the vignette](https://cran.r-project.org/web/packages/ggiraph/vignettes/shiny_usage.html).